---
title: "HOMEWORK 1 - PSTAT 220A"
author: "Jiajia Zheng"
date: "10/8/2017"
output:
   pdf_document: 
      latex_engine: xelatex
---
1. Write a R function that returns the following dispersion measures:
(a) Standard deviation;
(b) Mean absolute deviation n
(c) Median absolute deviation Median|xi − Median(x)|;
(d) Inter-quartile range

```{r}
dis.measures <- function(x){
  sd = sd(x)
  aad = mad(x, constant = 1)
  mad = mad(x)
  IQR = IQR(x)
  return(c(sd, aad, mad, IQR))
}
```
<\br>

<\br>

2. For the data set pima in library faraway, apply the R function you have constructed
above to compute standard deviations, mean absolute deviation, median absolute deviation
and inter-quartile range for the following variables: pregnant, diastolic, bmi and
age. Construct box-plots, histograms, QQ-plots and kernel density estimates for these
variables. Comment on features such as the distribution and outliers in these plots.

```{r}
library(MASS)
data(Pima.tr)
p <- Pima.tr
para_npreg <- dis.measures(p$npreg)
para_dias <- dis.measures(p$bp)
para_bmi <- dis.measures(p$bmi)
para_age <- dis.measures(p$age)
```

Start plotting the parameters:
```{r}
par(mfrow = c(2,2))
boxplot(p$npreg, ylab = "Pregnance Times")
hist(p$npreg, font.main = 1)
qqnorm(p$npreg, font.main = 1)
plot(density(p$npreg), font.main = 1)
mtext("Plots of Variable: Pregnancy", outer = TRUE, line = -1, font = 2, side = 3)
```

Comments:

* The distribution is left skewed

* The median pregancy times is 2

* Only 25% women have more than 6 times of pregnancy, with 1 outlier  
  
<\br>  
  
```{r}
par(mfrow = c(2,2))
boxplot(p$bp, ylab = "Blood Pressure")
hist(p$bp, font.main = 1)
qqnorm(p$bp, font.main = 1)
plot(density(p$bp), font.main = 1)
mtext("Plots of Variable: Diastolic", outer = TRUE, line = -1, font = 2, side = 3)
```

Comments:

* The blood pressure accords to approximately normal distribution

* The median of diastolic blood pressure is 50

* There are several outliers  

<\br>

```{r}
par(mfrow = c(2,2))
boxplot(p$bmi, ylab = "BMI Value")
hist(p$bmi, font.main = 1)
qqnorm(p$bmi, font.main = 1)
plot(density(p$bmi), font.main = 1)
mtext("Plots of Variable: BMI", outer = TRUE, line = -1, font = 2, side = 3)
```

Comments:

* The BMI accords to approximately normal distribution

* The median of BMI is 33, with no outliers  

<\br>
```{r}
par(mfrow = c(2,2))
boxplot(p$age, ylab = "Age")
hist(p$age, font.main = 1)
qqnorm(p$age, font.main = 1)
plot(density(p$age), font.main = 1)
mtext("Plots of Variable: Age", outer = TRUE, line = -1, font = 2, side = 3)
```

Comments:

* The age distribution is highly left skewed

* The median age is 28, with no outliers

* Only 25% women were over 40 years old  

<\br>

<\br>

3. The following observations are precipitation values during March over a 30-year period
in Minneapolis-St. Paul.

       0.77 1.20 3.00 1.62 2.81 2.48 1.74 0.47 3.09 1.31
       1.87 0.96 0.81 1.43 1.51 0.32 1.18 1.89 1.20 3.37
       2.10 0.59 1.35 0.90 1.95 2.20 0.52 0.81 4.75 2.05

(a) Construct QQ-plot. Is normality plausible?
(b) If not, can you find a transformation such that the normality assumption is appropriate?

```{r}
install.packages("xlsx", repos = "http://cran.us.r-project.org")
library(readxl)
MS_rain <- read_excel("~/Documents/Precipitation.xlsx")
qqnorm(MS_rain$Precipitation); qqline(MS_rain$Precipitation)
```

The above qqplot shows that it does not apply to normal distribution.  


```{r}
# So let's try lognormal distribution

MS_rain["ln_MS_rain"] <- log(MS_rain$Precipitation)
qqnorm(MS_rain$ln_MS_rain); qqline(MS_rain$ln_MS_rain)
```

The log value of precipitation applies to normal distribution.    
  
<\br>

<\br>

4. The following data are failure times in hours of 45 transmissions from caterpillar tractors
belonging to a particular American company:

        4381 3953 2603 2320 1161 3286 6914 4007 3168
        2376 7498 3923 9460 4525 2168 1288 5085 2217
        6922 218 1309 1875 1023 1697 1038 3699 6142
        4732 3330 4159 2537 3814 2157 7683 5539 4839
        6052 2420 5556 309 1295 3266 6679 1711 5931

Use QQ-plots to examine the applicability of the following models for the probability
distribution of failure time: normal, lognormal, exponential and Gamma (hint: check
out the function fitdistr in the library MASS to fit these distributions. You may
rescale the data if you have numerical problems). For the model that fits best (explain
how you determine which model fits best), compare the PDF with the kernel density
estimate.


```{r}
library(readxl)

Failuretimes <- read_excel("~/Documents/Failuretimes.xlsx")
f <- Failuretimes$Failure
par(mfrow = c(1,1))
hist(f)
```

Now use *fitdistr* function:
```{r}
fit_norm <- fitdistr(f, "normal")
fit_norm$estimate
fit_norm$loglik
```
```{r}
fit_lognorm <- fitdistr(f, "lognormal")
fit_lognorm$estimate
fit_lognorm$loglik
```

```{r}
fit_exp <- fitdistr(f, "exponential")
fit_exp$estimate
fit_exp$loglik
```

```{r}
fit_gam <- fitdistr(f, "gamma", lower=0.0011)
fit_gam$estimate
fit_gam$loglik
```

Then draw qqlot of each distribution:
```{r}
par(mfrow = c(2,2))

qqplot(f, rnorm(1000, fit_norm$estimate[1], fit_norm$estimate[2]), ylab = "", main = "norm")
abline(0, 1)

qqplot(f, rlnorm(1000, fit_lognorm$estimate[1], fit_lognorm$estimate[2]), ylab = "", main = "lognorm")
abline(0, 1)

qqplot(f, rexp(1000, fit_exp$estimate[1]), ylab = "", main = "exp")
abline(0, 1)

qqplot(f, rgamma(1000, fit_gam$estimate[1], fit_gam$estimate[2]), ylab = "", main = "gamma")
abline(0, 1)
```

Normal distribution fits best based on the following qqplot results.  
  
<\br>

```{r}
# Let's compare the PDF with the kernel density estimate
fit_base <- seq(0, 10000, 1)
hist(f, freq = F)
lines(density(f))
lines(fit_base, dnorm(fit_base, fit_norm$estimate[1], fit_norm$estimate[2]), col = "red")
legend("topright", c("density estimate", "normal distribution"), col = c("black","red"), lty = c(1,1))
```

<\br>

<\br>
  
5. Generate 5,000 random samples from the exponential distribution with rate=1. Plot
the histogram and comment on the shape of the distribution. Now break 5,000 samples
into 50 groups with 100 samples in each group. Compute averages in each group and
plot the histogram of these averages. Do these two histograms have the same shape?
Why or why not?

```{r}
x <- rexp(5000, rate = 1)
hist(x)
```

The shape is higly left skewed with exponential distribution.  

<\br>

```{r}
# Now split the dataset into 50 groups with 100 samples in each group and calculate the mean
x_split <- as.data.frame(split(x, 1:50))
x_split_mean <- colMeans(x_split)
hist(x_split_mean)
```

The histogram of the subgroups' mean distribution shape is more like normal distribution, which is very different with the total group distribution.  
  
<\br>

<\br>

6. Generate 600 random samples from the normal distribution with mean 10 and standard
deviation 5. Divide these 600 samples into 100 groups each with 6 samples. Compute
the statistic (X¯ − 10)/sqrt(S^2/6) for each group. What kind of distribution do you expect this statistic to follow? Now you have 100 such statistics, can you check if these
statistics actually follow the expected distribution?


```{r}
install.packages("matrixStats", repos = "http://cran.us.r-project.org")
library(matrixStats)

a <- rnorm(600, mean = 10, sd = 5)
a_split <- as.data.frame(split(a, 1:100))
a_split_stat <- (colMeans(a_split) - 10)/sqrt((colSds(as.matrix(a_split))^2 / 6))

hist(a_split_stat)
```

```{r}
qqnorm(a_split_stat); qqline(a_split_stat)
```

The calculation results distribution is nearly normal distribution with only several outliers.
